Untitled

### 2022-12-06 Manuscript/trial.Rmd
figure 3 update

---
title: "Bee brain size is positively associated with urban lifestyles but not with positive population trends"
output:
  pdf_document:
    fig_caption: yes

link-citations: yes
linkcolor: RoyalBlue
urlcolor: RoyalBlue
fontsize: 12pt

bibliography: 
  - references.bib

header-includes:
- \usepackage{orcidlink}
- \usepackage{caption,setspace}
- \captionsetup[figure]{labelformat=empty}
- \usepackage{float} 
- \floatplacement{figure}{H} 
---

\captionsetup[table]{labelformat=empty}

```{r echo=FALSE,warning=FALSE}

knitr::opts_chunk$set(fig.pos = "H", out.extra = "")
```

\textbf{Miguel Á. Collado$^{1}$} \orcidlink{0000-0002-4216-317X} \textbf{Jose B. Lanuza$^{2}$} \orcidlink{0000-0002-0287-409X} \textbf{and Ignasi Bartomeus$^{2}$} \orcidlink{0000-0001-7893-4389}

Departamento de Ciencias de la Computación e Inteligencia Artificial, Universidad de Sevilla, Seville, Spain

Estación Biológica de Doñana (EBD-CSIC), E-41092 Seville, Spain

\textbf{Author for correspondence:}

# Abstract

The rapid conversion of natural habitats to anthropogenic landscapes is threatening insect pollinators worldwide, raising concern on the consequences for their fundamental role in ecosystems for fruit and seed production. However, not all pollinators are negatively affected by habitat conversion, but some find in anthropogenic landscapes appropriate resources to survive and reproduce. Why some animals can thrive in anthropogenic environments that most species cannot tolerate remains insufficiently understood, but it might in part reflect enhanced behavioural plasticity to exploit new resource opportunities. While this hypothesis has received ample support in studies of vertebrates, it is less obvious whether the small brain of insects can offer enough behavioural plasticity to cope with the many challenges of anthropogenic habitats. To address this issue, we measured brains for 145 individuals from 44 species of bees ---the most important group of pollinators--- from North America and Europe, and compared them with detailed information on habitat preferences and population trends. Our analyses revealed that bees that prefer forest habitats had smaller brains relative to their body size, but those who prefer urban habitats tended to have larger brains compared to forest dwellers. However, we found no evidence that either larger brains or a preference for urban habitats are associated with positive population trends. These results suggest that a large brain can help maintaining bee urban populations, but also highlight that being tolerant to urbanisation is not enough to compensate for the worldwide decline of their populations.

Keywords: encephalisation, brain-body mass ratio, habitat preference, Apoidea, GBIF

\clearpage

# Introduction

Relevant refs:

<https://onlinelibrary.wiley.com/doi/10.1111/ele.14133?af=R>

This may be also relevant:

Although recent studies challenge the use of absolute brain size as an estimator for behavioral complexity and computational power (Chittka and Niven, 2009, and references therein), many studies have shown correlations between brain size and performance (Deaner et al., 2007; Gronenberg and Couvillon, 2010; Lefebvre and Sol, 2008).

Pollination is a fundamental ecosystem service that connects the earth's vegetation and human economy. Pollination services are performed by both managed and wild pollinators, mostly bees [@james2008; @winfree2011], both of which might be affected by a range of current and projected human-related environmental changes. There are indeed clear shreds of evidence of recent declines in pollinator populations [@bartomeus2019], and parallel declines in the plants that rely upon them [@potts2010; @scheper2014]. For example, in the UK, the third part of wild pollinator species has decreased from 1980 to 2013 [@powney2019], including six of all 16 non-parasitic bumblebees [@williams2009]. The decline of pollinators currently observed is projected to increase in the future [@dormann2008]. Because most wild plant species (80%) and crop species (75%) are directly dependent on insect pollination for fruit and seed production [@klein2007; @ollerton2011], there is an increasing concern that the loss of pollinators from both natural communities and agricultural areas will have severe negative consequences for ecosystem functions and services [@ashman2004; @klein2007].

Although the reason why many pollinator are declining is insuficiently understood, a factor that is considered central is the loss and alteration of their habitats [@brown2009]. The destruction and fragmentation of natural habitats and their replacement by anthropogenic landscapes present new challenges for survival and reproduction of organisms, increasing the risk of extinction by negative population growth [@brown2009]. However, not all pollinators species are equally affected by habitat loss and alteration. In bees, for example, there are also "winners" ---like Bombus terrestris [@rasmont2008]--- that are able to thrive in human-altered environments, and there are also other species that although do not proliferate in such environments can at least maintain self-sustaining populations [@biesmeijer2006; @bartomeus2013; Reemer et al., 2012????; @collado2019].

Human-dominated habitats ---notably cities--- offer unique ecological opportunities in the form of new nesting spots, reduced predation pressure and high food availability associated with non-indigenous plants [@cane2006; @winfree2011; @collado2019]. Although variation across species in tolerance to habitat loss and alteration is likely to have important demographic consequences, whether this is the reason why some pollinators are declining while others are increasing is currently unknown. Hence, the ability to adapt to human-dominated new habitats and to exploit new opportunities have become important for species to thrive [@manenti2013]. For species that tolerate human-related disturbances, human-dominated habitats ---notably cities--- offer unique ecological opportunities in the form of new nesting spots, reduced predation pressure and high food availability associated with non-indigenous plants [@cane2006; @winfree2011; @collado2019]. The question is then why only some species are able to enjoy the benefits of an urban life.In a novel environment, the chances to successfully survive and reproduce largely depends on cognition, that is, the processes involved in gathering, storing and reacting to environmental information. Cognitive limitations have thus help explaining the low success of many species in novel environments in general, and urban areas in particular. For example, reptiles, amphibians and birds that flourished after being introduced to new regions outside their native ranges tend to have larger brains (relative to their body size) than unsuccessful invaders [@sol2008; @sol2005; @amiel2011]. In vertebrates, relative brain size is a good proxy of problem-solving and learning (e.g. see chapter 2 of this thesis for its relationship with learning), which can help individuals to cope with the many challenges on novel environments. However, similar evidence is lacking for insects, in part because advanced cognitive abilities were not considered possible with their tiny brains. More important, there was no unequivocal evidence that brain size was related to cognitive performance. However, we now know that bees and other insects have advanced cognition, being able for instance to use tools to resolve new problems and learn socially how to do so [@alem2016]. There is in addition evidence that cognitive differences in the brain are reflected in differences in their brain size (Chapter 2). We used previously published data on bee forest and urban preference [@collado2019] as opposite points of habitat complexity, expecting to find bigger brain sizes for urban dwellers because a wider array of cognitive abilities are needed to inhabit these habitats. Species population trends were extracted from @bartomeus2013 and both datasets were crossed with a newly collected brain weight database. We wanted to explore if there is a relationship between brain size, habitat preference, and population trends. Hence, our questions were the following:Are brain sizes, as a proxy of behavioural plasticity, related to habitat preferences?Is population stability related to brain size as it has been seen in other taxa?Are habitat preferences related to population stability? Or alternatively, are urban dwellers populations increasing?

\clearpage

# Methods

## Overview

Our study explores the long standing idea that species with higher encephalisation (i.e., large brain mass relative to their body) are more likely to thrive in urbanised areas. To evaluate this hypothesis, we have used a dataset of bee species captured in North America and Europe with measurements of brain mass and body size. For all these species, we downloaded the occurrence records from the Global Biodiversity Information Facility (GBIF) and selected as study area the states or countries with highest density of records. After filtering the data to avoid possible biases, we extracted the land use classification of the different georeferenced records and summarised them into 3 different categories when possible: (1) natural, (2) agricultural and (3) urban. To evaluate if the different bee species occur more in some habitats than others, we calculated their habitat preference by comparing the observed occurrences with the ones of null models. Finally, for the species that showed low and high statistical preference for the different habitats, we investigated how the association between habitat preference and encephalisation changed by habitat type.

All analyses were undertaken in R version 4.0.5 [@r2021] and the manuscript was written in rmarkdown version 2.17 [@markdown2022].

```{r Taxonomic information about the dataset, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE}

library(readr)
library(dplyr)

#Load final dataset 
raw = read_csv("../Raw_data/all_cleaning.csv")

d = raw %>% 
group_by(Species) %>% 
summarise(individuals = n())

#Load taxonomic rank script
tax = read_csv("../Data/Processing/taxonomic_rank.csv")

## (a) Bee species dataset

Our dataset contains measurements of brain and body size for bee specimens captured by hand netting in different areas of the East Coast of United States and Europe (i.e., Spain and Netherlands). In total, the dataset includes `r sum(d$individuals)` female individuals from `r nlevels(factor(unique(d$Species)))` species that represent `r nlevels(factor(unique(tax$family)))` different families and `r nlevels(factor(unique(tax$genus)))` genera. We only considered female specimens because: (i) females are involved in a larger number of tasks than males and are likely to experience greater selective pressures from the environment; and (ii), there are evident behavioral and morphological differences between sexes that are associated with the size and structure of the brain [@streinzer2013; @roselino2015]. In addition, because brain size can change with age and experience of the different individuals within a species [@durst1994; @withers2008], we limited our sampling to specimens foraging on flowers. Note that 93 of the species used in our study were previously published with their respective trait measurements in @sayol2020.

## (b) Brain and body size measurements

Brain size was measured as the weight of fixed brain for each specimen. First, all collected individuals were kept in cold and sent them to the laboratory where they were identified by expert taxonomists and anesthetise in cold [@crook2013]. Then, heads were removed with a scalpel and placed in a fixative solution (4% paraformaldehyde in 0.1 M phosphate-buffered saline -PBS- pH = 7.4) to avoid degradation. Subsequently, brains were extracted from the head capsule and placed on a petri dish after removing the retina from the optic lobes and cleaning the entire brain of all tracheae and fat bodies. By doing this, we ensured that our brain measurements mainly accounted for neural tissue. We then placed the brains on a small piece of Parafilm® and dried the exceeding fixative solution with Kimwipes® tissues. Finally, within 4 seconds of the liquid removal, we weighted the brains with the help of a microbalance Sartorius Cubis® and calculated the average weight for the species with more than one specimen.

We considered as a proxy of body size the average intertegular span (ITS) per species. ITS represents the distance between the base of the wings (tegulae) on the bee thorax and has been shown to accurately predict body size as it is highly correlated with dry body weight [@cane1987; @kendall2019]. All ITS measurements were conducted with a stereomicroscope (magnification 16x or 80x) with a calibrated ocular micrometre (resolution down to 0.02 mm).

## (c) Occurrence data

```{r Loading study site areas, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE}

#Calculate study site areas
#Load libraries
library(sf)
library(sfheaders)

#EUROPE:
#Read data
euro_map <- st_read("../Data/Europe_data/euro_map_1.shp", quiet = TRUE)
#Add area col
euro_map$area <- st_area(euro_map)
#Create dataframe with the correct units
d1 = euro_map %>% group_by(CNTR_CODE) %>% summarize(st_union(geometry), area_NAME = sum(area)) %>% 
mutate(areakm2 = units::set_units(area_NAME, value = "km^2"))
#Calculate area
eukm2 = prettyNum(as.numeric(round(sum(d1$areakm2),0)),big.mark=",")

#USA:
#Read data
usa_states_1 <- read.csv("../Data/Usa_data/usa_states.csv")
#Not an sf object, convert into multypolygon
usa = sfheaders::sf_multipolygon(
  obj = usa_states_1
  , multipolygon_id = "group"
  , x = "long"
  , y = "lat"
) %>% st_set_crs(4326)
#Calculate area
usa$area <- st_area(usa)
#Create dataframe with the correct units
d2 = usa %>% group_by(group) %>% summarize(st_union(geometry), area_NAME = sum(area)) %>% 
mutate(areakm2 = units::set_units(area_NAME, value = "km^2"))
#Sum areas
usakm2 = prettyNum(as.numeric(round(sum(d2$areakm2),0)),big.mark=",")

```{r Species and occurrences info, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE}

#Read occurrence records and extract relevant info
#Load libraries
library(data.table)
#Load clean data with bee occurrences
europe <- read.table("../Data/Europe_data/all_above_50_europe.csv.gz",  header=T, quote="\"", sep=",")
usa <- read.table("../Data/Usa_data/all_above_50_usa.csv.gz",  header=T, quote="\"", sep=",")

#Europe
eu_occurrences = prettyNum(nrow(europe), big.mark=",")
records_eu = europe %>% 
group_by(species) %>% 
summarize(no_rows= length(species))

#USA
usa_occurrences = prettyNum(nrow(usa), big.mark=",")
records_usa = usa %>% 
group_by(species) %>% 
summarize(no_rows= length(species))

#Number of species
usa_spp = nrow(records_usa)
eu_spp = nrow(records_eu)
spp_number = nrow(records_eu) + nrow(records_usa)
#Median number of records per dataset
eu_median = median(records_eu$no_rows)
usa_median = median(records_usa$no_rows)

#Load now dataset with full set of species
all_spp = readRDS("../Data/Processing/wit.mean.rds")
spp_number_nofilter = nrow(all_spp) #total number of species

We downloaded occurrence information for all species from the Global Biodiversity Information Facility (GBIF; http://www.gbif.org/) for North America and Europe. The data was downloaded through R programming language with the help of the function *occ_download* from the package *rgbif* version 3.7.3 [@chamberlain2017]. Because the records were biased towards some areas that were associated with political boundaries, we selected the states or countries with highest density of records for our set of species. We selected the boundaries from the maps packages *spData* version 2.2.0 [@spdata] and *giscoR* vesion 0.3.2 [@giscoR], and extracted the occurrences with the help of the function *st_intersection* from the package *sf* version 1.0-7 [@pebesma2018]. For North America, we selected states located on the East Coast of United States (Figure 1A) that covered an approximate area of `r usakm2` km$^{2}$. These states were: Vermont, New Hampshire, Massachusetts, Rhode Island, Connecticut, New York metropolitan area, New Jersey, Delaware and Maryland. Similarly, we selected countries located on the north and center of Europe (Figure 1B) that represented a total area of `r eukm2` km$^{2}$. These countries were: England, Belgium, Netherlands, Germany and Denmark. All occurrence data used in this study can be downloaded from https://doi.org/10.15468/dl.nenuem where just the countries of interest were retrieved. To avoid biases in our analysis, we filtered the occurrence records according the following criteria: (i) we only included species with a widespread distribution of occurrences along our study area; (ii) all species had a minimum number of 50 records; (iii) all duplicated records were excluded (i.e., records of specimens with the same date and location); (iv) we only included georeferenced records with a minimum of two decimals of latitude/longitude coordinates; and (v), we only included records between the time interval 1990 and 2022 in order to avoid large differences with the extracted land cover classifications from 2006. After pruning the data, our final dataset comprised `r spp_number` species with  `r usa_occurrences` occurrence records for United States (`r usa_spp` species; median occurrences = `r usa_median`) and `r eu_occurrences` records for Europe (`r eu_spp` species; median occurrences = `r eu_median`).

```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, fig.align="center", fig.height=12, fig.width=12, fig.cap="\\textbf{Figure 1.} Maps showing the study areas and the number of GBIF occurrences in logarithmic scale for the selected bee species. A) East Coast of United States with records for Vermont, New Hampshire, Massachusetts, Rhode Island, Connecticut, New York metropolitan area, New Jersey, Delaware and Maryland. B) European continent with records downloaded for England, Belgium, Netherlands, Germany and Denmark.",  out.width="90%"}

#Plot maps with density of bee occurrences

#Europe----
#Load libraries
library(giscoR) #map of europe
library(ggplot2)
library(dplyr)
library(readr)
library(sf)
library(patchwork)
library(rnaturalearth)

#Load europe map
Europe <- gisco_get_countries(region = "Europe")
#Load countries filtered
euro_map <- st_read("../Data/Europe_data/euro_map_1.shp", quiet = TRUE)
#Load data with records filtered
all_long_lat <- read.table("../Data/Europe_data/all_above_50_europe.csv.gz",  header=T, quote="\"", sep=",")

#Plot Europe map with counties of interest highlighted
p1 = ggplot(Europe) +
geom_sf(fill="gray25",color = "black", alpha = 0.5,size=0.2)  +
geom_sf(data= euro_map %>% group_by(CNTR_CODE) %>% summarise() ,
aes(fill = CNTR_CODE, group=CNTR_CODE),fill = "#f2ea69", color = "black",alpha=0.8, size = 0.25) +
coord_sf(xlim = c(-10, 30), ylim = c(35, 65))  + 
microplot::theme_collapse(panel.border = element_rect(colour = "black", fill=NA, size=2),
panel.background = element_rect(fill = "white"))

p2 = ggplot(Europe) +
geom_sf(fill="gray80",color = "black", alpha = 1,size=0.2)+
geom_bin2d(data = all_long_lat, aes(x = long, y = lat, fill=log(..count..)),size = 1,bins=70) +
viridis::scale_fill_viridis(name = "log(Count)", breaks = c(0, 4, 8), limits=c(0,8)) +
coord_sf(xlim = c(-8, 18), ylim = c(40, 60)) +
theme(panel.background = element_rect(fill = "white"),
panel.border = element_rect(colour = "black", fill=NA, size=1)) +
ylab("Latitude") +
xlab("Longitude") +
ggtitle("B") +
theme(legend.position = c(0, 1),legend.justification = c(0, 1),
legend.background = element_blank(),
legend.box.background = element_blank(),
legend.key = element_blank(), 
legend.title = element_text(face = "bold"),
axis.title = element_text(face="bold", size = 14),
plot.title = element_text(face="bold", size = 18),
legend.text = element_text(face = "bold", size = 12), legend.direction="horizontal") +
scale_y_continuous(breaks = seq(40, 60, by = 6)) +
scale_x_continuous(breaks = seq(-8, 18, by = 8)) +
guides(fill = guide_colourbar(title.position = "top", title.hjust = .5)) +
ggspatial::annotation_scale(location = 'br',  bar_cols = c("grey60", "white"),  width_hint = 0.15)

#Insert general plot within specific plot
panel1 = p2 + inset_element(p1, left = 0, bottom = -0.89 , right = 0.35, top = 1.23, ignore_tag = TRUE)

#USA-----
#Load libraries

#Load data
usa_states_1 <- read.csv("../Data/Usa_data/usa_states.csv")
ny_metropolitan_1 <- read.csv("../Data/Usa_data/ny_metropolitan.csv")
all_long_lat <- read.table("../Data/Usa_data/all_above_50_usa.csv.gz",  header=T, quote="\"", sep=",")

#Load maps
canada <- ne_states(country = "canada", returnclass = "sf") %>%
st_as_sf(crs = 4326)

usa <- ne_states(country = "united states of america", returnclass = "sf") %>%
st_as_sf(crs = 4326)

p1 = ggplot(canada) + 
geom_sf(fill="gray60",color = "black", alpha = 1,size=0.2) + 
geom_sf(data= usa, fill="gray60",color = "black", alpha = 1,size=0.2) + 
microplot::theme_collapse(panel.border = element_rect(colour = "black", fill=NA, size=2),
panel.background = element_rect(fill = "white")) +
geom_map(data = ny_metropolitan_1, 
map = ny_metropolitan_1,aes(long, lat, map_id = region, group = group), fill = "#f2ea69",color = "black",
size = 0.1) +
geom_map(data = usa_states_1, map = usa_states_1,
aes(long, lat, map_id = region, group = group), color = "black", fill = "#f2ea69",
size = 0.1) +
coord_sf(xlim = c(-86, -62), ylim = c(25, 48))

p2 = ggplot(canada) + 
geom_sf(fill="gray80",color = "black", alpha = 1,size=0.2) + 
geom_sf(data= usa, fill="gray80",color = "black", alpha = 1,size=0.2) +
geom_bin2d(data = all_long_lat, aes(x = long, y = lat, fill=log(..count..)),size = 1,bins=50) +
viridis::scale_fill_viridis(name = "log(Count)", breaks = c(0, 4, 8), limits=c(0,8)) +
coord_sf(xlim = c(-82, -67), ylim = c(33, 47))  +
theme(panel.background = element_rect(fill = "white"),
panel.border = element_rect(colour = "black", fill=NA, size=1)) +
ylab("Latitude") +
xlab("Longitude") +
ggtitle("A") +
theme(legend.position = c(0, 1),legend.justification = c(0, 1),
legend.background = element_blank(),
legend.box.background = element_blank(),
legend.key = element_blank(),
legend.title = element_text(face = "bold"),
axis.title = element_text(face="bold", size = 14),
plot.title = element_text(face="bold", size = 18),
legend.text = element_text(face = "bold", size = 12, hjust=0), legend.direction="horizontal") +
scale_y_continuous(breaks = seq(33, 47, by = 4)) +
scale_x_continuous(breaks = seq(-82, -67, by = 4)) +
guides(fill = guide_colourbar(title.position = "top", title.hjust = .5))+ 
ggspatial::annotation_scale(location = 'br',  bar_cols = c("grey60", "white"))

#Insert general plot within specific plot
panel2 = p2 + inset_element(p1, left = 0, bottom = -0.86, right = 0.35, top = 1.2, ignore_tag = TRUE)

library(patchwork)
panel2  + panel1

## (d) Land cover classification

We assigned a habitat type for all GBIF occurrences by merging land cover information with the different georeferenced records. The land cover classification was obtained from the 2006 online inventories of the National Land Cover Database (NLCD) for United States and the Corine Land Cover (CLC) for Europe. After downloading these as raster files, we used the functions *rast* and *extract* from the *Terra* package version 1.6-41 [@hijmans2022] to read and obtain the cover classification of the different georeferenced records, respectively. Then, to simplify the interpretation and conduct a joint analysis for both occurrence datasets, we divided the resulting cover classes into three single categories: (i) natural, (ii) agricultural and (iii) urban. The original extracted cover classes and our classification into these three different categories are shown in Table S1 for United States and Table S2 for Europe.

## (e) Construction of the phylogenetic tree

We constructed a phylogeny with all studied species by using a previously published genus-level bee phylogeny [@hedtke2013]. For this, we first read the backbone phylogeny with the function *read.tree* and drop the missing genera with function *drop.tip* from the package *ape* version 5.6-2 [@paradis2019]. Then, we selected as an outgroup a wasp genera (*Tachysphex*) and recalculated distances to the root with the function *chronos* from *ape*. Species tips were added to genera nodes as polytomies of equal branch length relative to the genera branch length using the phytools package version 1.2-0 [@revell2012]. From the resulting tree, we created a scaled phylogenetic covariance matrix with the function *inverseA* from the package *MCMCglmm* version 2.33 [@hadfield2010].

## (f) Analysis

We estimated brain size relative to the body (i.e., encephalisation) by modeling the log-transformed association between brain weight and intertegular span. For this, we specified a simple Bayesian regression with the function *brm* from the package *brms* version 2.18 [@burkner2017] where brain weight was the response variable and intertegular span the predictor. To correct for the non-independent evolutionary histories among species, we included the phylogenetic covariance matrix as random factor. Note that this matrix contained all species of our dataset (N = `r spp_number_nofilter`) because we expect better predictive power with a larger sample size than the one with the filtered dataset (N = `r spp_number`). Lastly, we extracted the differences between the observed and predicted values (i.e., residuals) that represent the degree of encephalisation. High values indicate larger brains than expected for their body size and low values denote smaller brains than expected.

Habitat preference was estimated by comparing species occurrences per habitat with the ones that we would expect if they occur at random. For this, we generated from our occurrence matrix 10,000 randomised matrices with help of the function *nullmodel* from the package *bipartite* version 2.16 [@dormann2008]. We implemented this function with the method 'r2dtable' which maintains row and column sums constant by using Patefield's algorithm [@patefield1981]. In other words, it maintains the proportional dominance of the species and habitats constant but reshuffles their associations. We then estimated the percentage of simulated occurrences per species and habitat that were under the observed ones. Finally, we considered that a bee species exhibited statistical preference when the observed occurrences per species and habitat were under the 5th percentile (low preference) or above the 95th percentile (high preference) in comparison with the simulated ones.

To evaluate if the degree of encephalisation changes by habitat type, we modeled with Bayesian generalised linear models their association. For this, we first joined the resulting habitat preference and encephalisation datasets by species and extracted those with statistical habitat avoidance or preference on any of these three habitats. Because we assessed habitat preference as binary (low or high), we specified a Bernouilli distribution where habitat preference was the response variable and encephalisation the predictor. Again, we included the phylogenetic covariance matrix as random factor. Moreover, we also explored how habitat preference changed by brain weight and intertegular span independently. For this, we conducted two analogous models with these two different predictors. Finally, we also investigated the different trends for United States and Europe separately.

All our models were run with 4,000 iterations with previous 1,000 warm up iterations and with non or very weakly informative priors from the brm function so the priors had negligible influence on the results [@burkner2017]. Finally, the different posterior predictive checks were conducted with the function *pp_check* also from the *brms* package.

# Results

```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=12, fig.width=12,fig.cap="\\textbf{Figure 2.} Phylogenetic relationship and habitat preferences for the selected bee species with brain weight and intertegular distance information (N=100). A) Phylogenetic tree at the genera level. Tree branches are coloured based on the geographical location of the different bee species (North of Europe, East Coast of United States or from both regions). The deviation of the brain size in relation to the body (i.e., residuals) is indicated with weighted points at the end of the tip branches. Larger points indicate larger brains in proportion to their body size and vice versa. B) Heatmap showing the habitat preference for each bee species. The columns delimit the habitat type (i.e., natural, agricultural, semideveloped and urban) and the rows the different bee species."}

library(ggtree) #v3.5.3.002
library(ggplot2)
library(stringr)
library(tidyverse)
library(tibble)

#Load phylogenetic data from europe
bee.tree100 = readRDS("../Data/Europe_USA/bee.tree100.rds")
#Load brain data
brain_weight = readr::read_csv("../Data/Processing/brain_weight_data.csv")
#Load preferences data
preferences_usa = readr::read_csv("../Data/USA_data/preferences_usa.csv") 
preferences_eu = readr::read_csv("../Data/Europe_data/preferences_europe.csv") 
#Bind rows
pref = bind_rows(preferences_usa, preferences_eu)

#Average preferences between USA and Europe
preferences = pref %>% 
group_by(Species) %>%
summarise(across(everything(), list(mean),.names = "{.col}")) %>% 
ungroup()

#Delete underscore from tree
bee.tree100$tip.label = str_replace(bee.tree100$tip.label, "_", " ")

#Generate vector with species
v = bee.tree100$tip.label
#Filter species with that are in the phylo
brain_weight = brain_weight %>% filter(Species %in% v)

#Fill a column based on condition
brain_weight = brain_weight %>% 
    mutate(Country = case_when((brain_weight$Species %in% preferences_eu$Species & 
    brain_weight$Species %in% preferences_usa$Species ~ "USA and Europe"),
    (brain_weight$Species %in% preferences_usa$Species ~ "USA"),
    (brain_weight$Species %in% preferences_eu$Species ~ "Europe")))

#Plot tree
p <- ggtree(bee.tree100) %<+% brain_weight  +  aes(color=Country)+
    geom_tippoint(aes(size = residuals, color=Country))+
    theme_tree2()+scale_size_continuous(range = c(-1, 3.6),name="Encephalisation")+
    scale_color_manual(name = "Data", values=c("darkgoldenrod1","cyan4","black"),
                       labels = c("Europe", "USA", "USA and Europe"),
                       breaks = c("Europe", "USA", "USA and Europe")) +
  ggtitle("A") + theme(plot.title = element_text(size=18, face="bold"))

#Convert wide format to long for plotting heatmap
d=gather(preferences, condition, measurement, Agricultural:Urban, factor_key=TRUE)

#Relevel for plotting
d = d %>% 
mutate(condition=fct_relevel(condition,c("Natural","Agricultural", "Urban")))

p2 = ggplot(d, aes(x=condition, y=Species)) + 
    geom_tile(aes(fill=measurement),color = "white") +
   # scale_fill_gradient(high="black",  low="white") +
    scale_fill_viridis_c(name = "Habitat\n preference") + 
    theme_minimal() +
    xlab(NULL) + 
    ylab(NULL) + 
    theme(axis.text.x = element_text(face="bold", color="black", 
    size=7.5,  vjust=1.15),
    axis.text.y = element_text(face="italic", color="black", size=7.5, hjust=0),
    plot.title = element_text(size=18, face="bold")) +
   ggtitle("B")

#Nice way to insert ordered heatmap
library(aplot)
p2 %>% insert_left(p) 
```

\clearpage

```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=5, fig.width=12, fig.cap="\\textbf{Figure 3.} Frequency distribution of the bee preferences for each habitat type (i.e., natural, agricultural and urban). The different habitats are reclassifications from the 2006 land use categories of the Corine Land Cover (CLC) inventory for Europe and the National Land Cover (NLC) database for Unites States.", fig.pos="p"}

@@

long_data$Preference_colour <- factor(long_data$Preference_colour, levels = c("A","C", "B"), labels = c("Low preference", "No preference", "High preference"))

ggplot(long_data, aes(x=Preference, color= Preference_colour)) +
    geom_histogram(aes(fill= Preference_colour),bins = 35, alpha = 0.8,linewidth=0.5) + 
    xlab("Habitat preference") +
    ylab("Frequency") +
    facet_wrap(~ Habitat) +
    scale_fill_manual(name="" ,values=c("coral2", "gray", "palegreen3")) +
    scale_color_manual(name="" ,values=c("coral2", "gray", "palegreen3")) +
    theme_bw() +   
    theme(axis.title = element_text(face = "bold", size = 14),
    panel.border=element_rect(size=1),
    strip.text = element_text(face = "bold", size = 18,hjust = -0.01),
    strip.background = element_rect(fill = NA, colour = NA, size = 0,linetype="solid"),
    panel.grid.minor=element_blank())+
    scale_y_continuous(expand = c(0, 0), limits = c(0,60))+
    geom_vline(aes(xintercept = 0.20), colour="black", linetype="dashed") +
    geom_vline(aes(xintercept = 0.80), colour="black", linetype="dashed") +
    scale_x_continuous(breaks = seq(from = 0, to = 1, by = 0.2),
                       labels = c(0,0.2,0.4,0.6,0.8,1)) +
  theme(legend.position = 'bottom', legend.box.background = element_rect(colour = "black",size=1))

@@

\clearpage

```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=6, fig.width=10, fig.cap="\\textbf{Figure 4.}", fig.pos="p"}

#Load libraries
library(readr)
library(dplyr)
library(ggplot2)

#Load data for plotting
long_data <- read_csv("../Data/Europe_USA/data_preference_residuals_europe_usa_qualitative.csv")

model1 = read_csv("../Data/Europe_USA/model_output_preference_residuals_europe_usa_qualitative.csv")

model2 = read_csv("../Data/Europe_USA/model_output_preference_brain_weight_europe_usa_qualitative.csv")

model3 = read_csv("../Data/Europe_USA/model_output_preference_it_europe_usa_qualitative.csv")

#Relevel for plotting
long_data = long_data %>% 
    mutate(Habitat=fct_relevel(Habitat,c("Natural","Agricultural", "Urban"))) 
#Relevel for plotting
model1 = model1 %>% 
    mutate(Habitat=fct_relevel(Habitat,c("Natural","Agricultural", "Urban"))) 
#Relevel for plotting
model2 = model2 %>% 
    mutate(Habitat=fct_relevel(Habitat,c("Natural","Agricultural", "Urban"))) 
#Relevel for plotting
model3 = model3 %>% 
    mutate(Habitat=fct_relevel(Habitat,c("Natural","Agricultural", "Urban")))

#colorBlindness::displayAllColors(viridis::viridis(4))

all_1 = ggplot(model1, aes(x = residuals, y = estimate__, color=Habitat)) +
geom_point(data =  long_data, aes(x = residuals, y = (Preference)), shape=21, size=2.5) +
geom_line(aes(color=Habitat),size=0.7) +
geom_ribbon(aes(ymin = lower__, ymax = upper__,  fill = Habitat),color=NA, linetype = "dashed",alpha = 0.2)+
theme_bw() +
ylab("Habitat preference") +
xlab("Encephalisation") +
scale_colour_viridis_d() +
scale_fill_viridis_d() +
scale_y_continuous(breaks = seq(0, 1, by = 1),labels = c("Low", "High")) +
ggtitle("A") +
theme(plot.title = element_text(face = "bold", size=18))

all_2 = ggplot(model2, aes(x = Brain.weight, y = estimate__, color=Habitat)) +
geom_point(data =  long_data, aes(x = Brain.weight, y = (Preference)), shape=21, size=2.5) +
geom_line(aes(color=Habitat),size=0.7) +
geom_ribbon(aes(ymin = lower__, ymax = upper__,  fill = Habitat),color=NA, linetype = "dashed",alpha = 0.2)+
theme_bw() +
ylab(NULL) +
xlab("Brain weight (mg)") +
scale_colour_viridis_d() +
scale_fill_viridis_d() +
scale_y_continuous(breaks = seq(0, 1, by = 1),labels = c("Low", "High")) +
ggtitle("B") +
theme(plot.title = element_text(face = "bold", size=18))

all_3 = ggplot(model3, aes(x = IT, y = estimate__, color=Habitat)) +
geom_point(data =  long_data, aes(x = IT, y = (Preference)), shape=21, size=2.5) +
geom_line(aes(color=Habitat),size=0.7) +
geom_ribbon(aes(ymin = lower__, ymax = upper__, fill = Habitat),color=NA, linetype = "dashed",alpha = 0.2)+
theme_bw() +
ylab(NULL) +
xlab("Intertegular distance (mm)") +
scale_colour_viridis_d() +
scale_fill_viridis_d() +
scale_y_continuous(breaks = seq(0, 1, by = 1),labels = c("Low", "High")) +
ggtitle("C") +
theme(plot.title = element_text(face = "bold", size=18))

library(patchwork)

all_1 + (all_2/all_3) + plot_layout(guides = "collect") &
theme(legend.position = 'bottom', panel.border=element_rect(size=1),
      axis.title=element_text(face="bold"),
      legend.box.background = element_rect(colour = "black",size=1))

\clearpage

# Discussion

>>>>

---
title: "Untitled"
output:
  pdf_document: default
  html_document: default
date: "2022-12-06"
---
```{r, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE,fig.align="center", fig.height=4, fig.width=4, fig.cap="\\textbf{Figure 3.} Frequency distribution of the bee preferences for each habitat type (i.e., natural, agricultural and urban). The different habitats are reclassifications from the 2006 land use categories of the Corine Land Cover (CLC) inventory for Europe and the National Land Cover (NLC) database for Unites States.", fig.pos="p"}

@@

long_data$Preference_colour <- factor(long_data$Preference_colour, levels = c("A","C", "B"), labels = c("Avoidance", "No preference", "Preference"))

ggplot(long_data, aes(x=Preference, color= Preference_colour)) +
    geom_histogram(aes(fill= Preference_colour),bins = 35, alpha = 0.8,linewidth=0.5) + 
    xlab("Habitat preference") +
    ylab("Frequency") +
    facet_wrap(~ Habitat) +
    scale_fill_manual(name="" ,values=c("coral2", "gray", "palegreen3")) +
    scale_color_manual(name="" ,values=c("coral2", "gray", "palegreen3")) +
    theme_bw() +   
    theme(axis.title = element_text(face = "bold"),
    panel.border=element_rect(size=1),
    strip.text = element_text(face = "bold", size = 18,hjust = -0.01),
    strip.background = element_rect(fill = NA, colour = NA, size = 0,linetype="solid"),
    panel.grid.minor=element_blank())+
    scale_y_continuous(expand = c(0, 0), limits = c(0,60))+
    geom_vline(aes(xintercept = 0.20), colour="black", linetype="dashed") +
    geom_vline(aes(xintercept = 0.80), colour="black", linetype="dashed") +
    scale_x_continuous(breaks = seq(from = 0, to = 1, by = 0.2),
                       labels = c(0,0.2,0.4,0.6,0.8,1))

@@
